\LARGE \bf Exploiting Non-Slip Wall Contacts \\ to Position Two Particles Using The Same Control Input

### 2018-12-11 journalWallFriction/resubmitv03/Responsev03.tex
cleaned images, proofed test, and prepped letter

	Thank you for this opportunity!

	\Reviewer{\#1}\response{In the revised version most of my previous concerns have been
addressed.  I think with minor revisions the manuscript could be
suitable for publication. 
The remaining issue still comes back to whether the iterative procedure is able to reach all target configurations.  The new manuscript shows numerically that the accessible space approaches the entire space for a circular 2D domain.  The paper, however, is meant to also cover any convex domains.  Can the authors also prove that the accessible space approaches the entire domain for any convex domain?  Or at least other polygonal cases?   I see that on p4 in the last paragraph of section C, there is a statement that particles can be steered to arbitrary positions, but without an explanation of how that result is obtained, which should be included.  I also think that paragraph should be highlighted a little more by a section title (to be parallel with the development of the circular workspace).  Assuming this can be explained then I would be satisfied with the manuscript.
}
	Thank you! This was a fun extension to work on.  In a similar form to the work with circular domain we now show that the accessible space rapidly grows to include the entire $\Delta$ configuration space.  For finite-sided polygons, this convergence requires finite time (assuming the particles have at least $\epsilon$ separation).  We have augmented Fig.~4 to show (as needed) the 4, 6, and 8-move reachable sets for polygons up to 20 sides.  These sets reach the entire $\Delta$ configuration space.

@@

	``\revision{
As the corner angles increase, the number of moves required to access the entire configuration space increases.  
As shown in Fig.~4,  four moves are sufficient for 3-sided polygons, and six moves accesses the entire $\Delta$ configuration space for up to 8-sided polygons.  Eight moves is sufficient for 20-sided polygons, but we have not checked polygons with more sides. For a circular workspace, with corner angles of 180$^\circ$, the worst-case configuration can only be approached asymptotically, as explained in Section~VI.}''

>>>>

	Thank you for this opportunity! In addition to the changes listed below, we used this opportunity to remove a few typos and clean up several figures.

	\Reviewer{\#1}\response{In the revised version most of my previous concerns have been
addressed.  I think with minor revisions the manuscript could be
suitable for publication. 
The remaining issue still comes back to whether the iterative procedure is able to reach all target configurations.  The new manuscript shows numerically that the accessible space approaches the entire space for a circular 2D domain.  The paper, however, is meant to also cover any convex domains.  Can the authors also prove that the accessible space approaches the entire domain for any convex domain?  Or at least other polygonal cases?   I see that on p4 in the last paragraph of section C, there is a statement that particles can be steered to arbitrary positions, but without an explanation of how that result is obtained, which should be included.  I also think that paragraph should be highlighted a little more by a section title (to be parallel with the development of the circular workspace).  Assuming this can be explained then I would be satisfied with the manuscript.
}
	Thank you! This was a fun extension to work on.  In a similar form to the work with circular domain we now show that the accessible space rapidly grows to include the entire $\Delta$ configuration space.  For finite-sided polygons, this convergence requires finite time (assuming the particles have at least $\epsilon$ separation).  We have augmented Fig.~4 to show (as needed) the 4, 6, and 8-move reachable sets for polygons up to 10 sides.  These sets reach the entire $\Delta$ configuration space.

@@

	``\revision{
As the corner angles increase, the number of moves required to access the entire configuration space increases.  
As shown in Fig.~4,  four moves are sufficient for 3-sided regular polygons, and six moves accesses the entire $\Delta$ configuration space for up to 8-sided regular polygons.  Eight moves is sufficient for polygons with less than 14-sides, but we have not checked polygons with more sides. For a circular workspace, with corner angles of 180$^\circ$, the worst-case configuration can only be approached asymptotically, as explained in Section~VI.
If the polygon is irregular, more moves may be required.  The entire configuration space is reachable in four moves for an acute triangle, but obtuse triangles require six moves.
}''

### 2018-12-11 journalWallFriction/simulation.tex
cleaned images, proofed test, and prepped letter

 \subsection{Example: Shortest Path in a Square Workspace}\label{subsec:square}
 If the goal configuration cannot be reached in one move but can be reached in three moves, the shortest path has a simple solution. The first move, $m_1$, makes one particle contact a wall, $m_2$ adjusts the relative spacing error  to zero, and $m_3$ takes the particles to their final ositions. 
$m_2$ cannot be shortened, so optimization depends on choosing the location where the particle contacts the wall. 
 Since the shortest distance between two points is a straight line, reflecting the goal position across the boundary wall and plotting a straight line gives the optimal contact location, as shown in Fig. \ref{fig:reflection}. 
  There are four walls, and four candidate solutions, but some candidate solutions may be invalid because a different boundary is hit before the desired first contact position in move $m_1$ (light grey regions) or  invalid because $m_2$ cannot generate the goal relative spacing (dark grey regions).
 %
\begin{figure}
\centering
\begin{overpic}[width=\columnwidth]{Reflection.pdf}\end{overpic}
\vspace{-2em}
\caption{\label{fig:reflection} In a square workspace, the shortest three-move path that reconfigures two particles from starting positions 
 (\protect\tikz \protect\draw[red,fill=white,line width=0.3mm] (-1ex,0) rectangle (0,1ex);,
 \protect\tikz \protect\draw[blue,fill=white,line width=0.3mm] (-1ex,0) rectangle (0,1ex);)
 to goal positions  
 (\protect\tikz \protect\draw[red,fill=white,line width=0.3mm] (0,0) circle (.5ex);,
 \protect\tikz \protect\draw[blue,fill=white,line width=0.3mm] (0,0) circle (.5ex);) are fixed.
 has the property that the incident angle equals the reflected angle, as shown at left. (right) The first contact is colored red if the red particle is the first to touch a boundary, and colored blue if the blue particle is the first to touch.
} \vspace{-1em}
\end{figure}
%

@@

%  This equation can be simplified to 
% 
% \begin{equation}
%\sqrt{1+s^2-2 s \cos \theta}+  \sqrt{1+e^2-2 e \cos(\beta-\theta)}. 
% \end{equation}
% 
\begin{figure}
\centering
\renewcommand{\figwid}{\columnwidth}
{\begin{overpic}[width =\figwid]{shortestpath.pdf}
\end{overpic}
}
\caption{\label{fig:shortestpath}{The shortest path between two points $S$ to $E$ in the unit disk that intersects the circumference. The path length as a function of intersection point, $P= (\cos\theta,\sin\theta)$ is shown at right. See \cite{BeckerShortestPath}.}
%\vspace{-1em}
}
\end{figure}

@@

\begin{overpic}[width=0.67\columnwidth]{Move4.pdf}\put(55.5,32.5){\begin{turn}{50.5} 
{\begin{tikzpicture}[thick]
\draw [red,   -  , dotted      ] (1,14.0) -- (0,14.0);
\end{tikzpicture}}\end{turn}}\end{overpic}
\begin{overpic}[width=0.67\columnwidth]{Move5.pdf}\put(55.5,32.5){\begin{turn}{50.5} 
{\begin{tikzpicture}[thick]
\draw [red,   -   , dotted     ] (1,14.0) -- (0,14.0);
\end{tikzpicture}}\end{turn}}\end{overpic}
\begin{overpic}[width=0.67\columnwidth]{finalMove.pdf}\end{overpic}
\caption{\label{fig:reachableSet}
Frames from reconfiguring two particles. 
Top six images show a polygonal workspace and the corresponding $\Delta$ configuration space with its 2-move reachable sets. 
Bottom six images show a disk-shaped workspace and the corresponding $\Delta$ configuration space with its 2-move reachable sets. 
For each the move 1 and 3 are simple translations of both particles and so the reachable sets do not change.
The reachable set morphs during move 2 because one particle is held stationary by the boundary. 
See multimedia attachment for animations of each.
}
\end{figure*}

%\subsection{Position Control of Two Robots}
\begin{figure}
\centering
\begin{overpic}[width=0.49\columnwidth]{middlegoalnum.pdf}\put(0,75){a)}\end{overpic}
\begin{overpic}[width=0.49\columnwidth]{middlegoaldist.pdf}\put(0,75){b)}\end{overpic}
\begin{overpic}[width=0.49\columnwidth]{worstnum.pdf}\put(0,75){c)}\end{overpic}
\begin{overpic}[width=0.49\columnwidth]{worstdist.pdf}\put(0,75){d)}\end{overpic}
\caption{\label{fig:contour}
%Plots show performance with one goal on the boundary.
Contour plots showing the number of moves and distance commanded if red particle's goal position is varied in $x$ and $y$. 
Starting positions of red and blue particles  
 (\protect\tikz \protect\draw[red,fill=white,line width=0.3mm] (-1ex,0) rectangle (0,1ex);,
 \protect\tikz \protect\draw[blue,fill=white,line width=0.3mm] (-1ex,0) rectangle (0,1ex);)
 and goal position of blue particle  \protect\tikz \protect\draw[blue,fill=white,line width=0.3mm] (0,0) circle (.5ex); are fixed.
The top row has the blue particle's goal position  at the origin, generating symmetric contour plots.
Moving the  blue particles' goal position  to $(-0.2,0)$, generates non-symmetric contour plots.
}
\end{figure}

\begin{figure}
\centering
%\begin{overpic}[width=\columnwidth]{deltanum.pdf}\end{overpic}\\
%\vspace{1em}
\begin{overpic}[width=\columnwidth]{deltadist.pdf}
\put(65,22){\scriptsize
$\Delta s$
\protect\tikz \protect\draw[myDarkGreen,fill=myDarkGreen,line width=0.3mm] (-1ex,0) rectangle (0,1ex); $=$
\protect\tikz \protect\draw[red,fill=white,line width=0.3mm] (-1ex,0) rectangle (0,1ex); $-$
\protect\tikz \protect\draw[blue,fill=white,line width=0.3mm] (-1ex,0) rectangle (0,1ex);}
\end{overpic}
\vspace{-1em}
\caption{\label{fig:deltanumdist}
The worst-case path length occurs when particles must swap antipodes. This can never be achieved but can be asymptotically approached. Plot shows decreasing error (radius $\delta$ around goal positions) as the number of moves grows.
 Red fit line is $8.66/(\textrm{distance}^3)$, which has an R-squared value of 0.77.
} 
\end{figure}

\begin{figure}
\centering
\renewcommand{\figwid}{1\columnwidth}
{
\begin{overpic}[width =\figwid]{contourDistnew.png}\put(-2,10){\begin{turn}{90} \tiny{unique particles}
\end{turn}}
%
\end{overpic}
\vspace{1em}
\begin{overpic}[width =\figwid]{JustSimulationV6.png}\put(-2,10){\begin{turn}{90} \tiny{unique particles}
\end{turn}}
%
\end{overpic}
\begin{overpic}[width =\figwid]{identical.png}\put(-2,6){\begin{turn}{90} \tiny{interchangeable particles}
\end{turn}}
\end{overpic}
}\caption{\label{fig:contourPlots}{Starting positions of particles $1$ and $2$ 
 (\protect\tikz \protect\draw[blue,fill=white,line width=0.3mm] (-1ex,0) rectangle (0,1ex);,
 \protect\tikz \protect\draw[myMagenta,fill=white,line width=0.3mm] (-1ex,0) rectangle (0,1ex);)
and goal position of particle $2$ (\protect\tikz \protect\draw[myMagenta,fill=white,line width=0.3mm] (0,0) circle (.5ex);) are fixed, and $\epsilon=0.001$.
 The top row of contour plots show the distance if particle $1$'s goal position is varied in $x$ and $y$. The middle row shows the number of moves required for the same configurations. The bottom row shows the same configuration but when the particles are interchangeable.}
\vspace{-1em}
}
\end{figure}
Algorithm \ref{alg:optimalAlg}  was implemented in Mathematica using particles with zero radius. Figure \ref{fig:reachableSet} shows frames of the algorithm in two representative workspaces, square and disk, with two arbitrary starting and goal configurations.
%An online interactive demonstration and source code of the algorithm are available at \cite{Shahrokhi2015mathematicaParticle}.
%  Fig.~\ref{fig:shapeControlMathematica1}  shows  an implementation of this algorithm with robot initial positions represented by hollow squares and final positions by circles. 
 %Dashed lines show the shortest route if robots could be controlled independently, while solid lines show the optimal shortest  path using uniform inputs.

 The contour plots in Fig.~\ref{fig:contour} left show the length of the path for two different settings. The top row considers \{$s_1,s_2,g_1$\} = \{$(0.2,0.2),(-0.1,-0.1),(0,0)$\} and the bottom row considers  \{$s_1,s_2,g_1$\} = \{$(0.2,0.2),(-0.1,-0.1),(-0.2,0)$\}, each in a workspace with $r= 0.5$, and $g_2$ ranging over all the workspace. Fig.~\ref{fig:contour} right shows the number of moves and left shows the total distance of the path. This plot shows the nonlinear nature of the path planning. When the goal is in the middle of the workspace, a symmetry in the path length is expected as the top row shows. The bottom row shows a shift in the goal position which breaks the symmetry of the path length in the workspace.

>>>>

 \subsection{Example: Shortest Path in a Square Workspace}\label{subsec:square}
 If the goal configuration cannot be reached in one move but can be reached in three moves, the shortest path has a simple solution. The first move, $m_1$, makes one particle contact a wall, $m_2$ adjusts the relative spacing error  to zero, and $m_3$ takes the particles to their final positions. 
$m_2$ cannot be shortened, so optimization depends on choosing the location where the particle contacts the wall. 
 Since the shortest distance between two points is a straight line, reflecting the goal position across the boundary wall and plotting a straight line gives the optimal contact location, as shown in Fig. \ref{fig:reflection}. 
  There are four walls, and four candidate solutions, but some candidate solutions may be invalid because a different boundary is hit before the desired first contact position in move $m_1$ (light grey regions) or  invalid because $m_2$ cannot generate the goal relative spacing (dark grey regions).
 %
\begin{figure}
\centering
\begin{overpic}[width=\columnwidth]{Reflection.pdf}\end{overpic}
\vspace{-2em}
\caption{\label{fig:reflection} In a square workspace, the shortest three-move path that reconfigures two particles from starting positions 
 (\protect\tikz \protect\draw[red,fill=white,line width=0.3mm] (-1ex,0) rectangle (0,1ex);,
 \protect\tikz \protect\draw[blue,fill=white,line width=0.3mm] (-1ex,0) rectangle (0,1ex);)
 to goal positions  
 (\protect\tikz \protect\draw[red,fill=white,line width=0.3mm] (0,0) circle (.5ex);,
 \protect\tikz \protect\draw[blue,fill=white,line width=0.3mm] (0,0) circle (.5ex);) 
 has the property that the incident angle equals the reflected angle, as shown at left. (right) The first contact is colored red if the red particle is the first to touch a boundary, and colored blue if the blue particle is the first to touch.
} \vspace{-1em}
\end{figure}
%

@@

%  This equation can be simplified to 
% 
% \begin{equation}
%\sqrt{1+s^2-2 s \cos \theta}+  \sqrt{1+e^2-2 e \cos(\beta-\theta)}. 
% \end{equation}
% 
\begin{figure}
\centering
\renewcommand{\figwid}{\columnwidth}
{\begin{overpic}[width =\figwid]{shortestpath3.pdf}
\end{overpic}
}
\caption{\label{fig:shortestpath}{The shortest path between two points $S$ to $E$ in the unit disk that intersects the circumference. The path length as a function of intersection point, $P= (\cos\theta,\sin\theta)$ is shown at right. See \cite{BeckerShortestPath}.}
%\vspace{-1em}
}
\end{figure}

@@

\begin{overpic}[width=0.67\columnwidth]{Move4.pdf}\put(55.5,32.5){\begin{turn}{50.5} 
{\begin{tikzpicture}[thick]
\draw [red,   -  , dotted      ] (1,14.0) -- (0,14.0);
\end{tikzpicture}}\end{turn}}\end{overpic}
\begin{overpic}[width=0.67\columnwidth]{Move5.pdf}\put(55.5,32.5){\begin{turn}{50.5} 
{\begin{tikzpicture}[thick]
\draw [red,   -   , dotted     ] (1,14.0) -- (0,14.0);
\end{tikzpicture}}\end{turn}}\end{overpic}
\begin{overpic}[width=0.67\columnwidth]{finalMove.pdf}\end{overpic}
\caption{\label{fig:reachableSet}
Frames from reconfiguring two particles. 
Top six images show a polygonal workspace and the corresponding $\Delta$ configuration space with its 2-move reachable sets. 
Bottom six images show a disk-shaped workspace and the corresponding $\Delta$ configuration space with its 2-move reachable sets. 
For each, moves 1 and 3 are simple translations of both particles and so the reachable sets do not change.
The reachable set morphs during move 2 because one particle is held stationary by the boundary. 
See multimedia attachment for animations of each.
}
\end{figure*}

%\subsection{Position Control of Two Robots}
\begin{figure}
\centering
\begin{overpic}[width=0.49\columnwidth]{middlegoalnum.pdf}\put(0,75){a)}\end{overpic}
\begin{overpic}[width=0.49\columnwidth]{middlegoaldist.pdf}\put(0,75){b)}\end{overpic}
\begin{overpic}[width=0.49\columnwidth]{worstnum.pdf}\put(0,75){c)}\end{overpic}
\begin{overpic}[width=0.49\columnwidth]{worstdist.pdf}\put(0,75){d)}\end{overpic}
\caption{\label{fig:contour}
%Plots show performance with one goal on the boundary.
Circular workspace: contour plots showing the number of moves and distance commanded if red particle's goal position is varied in $x$ and $y$. 
Starting positions of red and blue particles  
 (\protect\tikz \protect\draw[red,fill=white,line width=0.3mm] (-1ex,0) rectangle (0,1ex);,
 \protect\tikz \protect\draw[blue,fill=white,line width=0.3mm] (-1ex,0) rectangle (0,1ex);)
 and goal position of blue particle  \protect\tikz \protect\draw[blue,fill=white,line width=0.3mm] (0,0) circle (.5ex); are fixed.
The top row has the blue particle's goal position  at the origin, generating symmetric contour plots.
Moving the  blue particles' goal position  to $(-0.2,0)$, generates non-symmetric contour plots.
}
\end{figure}

\begin{figure}
\centering
%\begin{overpic}[width=\columnwidth]{deltanum.pdf}\end{overpic}\\
%\vspace{1em}
\begin{overpic}[width=\columnwidth]{deltadist.pdf}
\put(65,22){\scriptsize
$\Delta s$
\protect\tikz \protect\draw[myDarkGreen,fill=myDarkGreen,line width=0.3mm] (-1ex,0) rectangle (0,1ex); $=$
\protect\tikz \protect\draw[red,fill=red,line width=0.3mm] (-1ex,0) rectangle (0,1ex); $-$
\protect\tikz \protect\draw[blue,fill=blue,line width=0.3mm] (-1ex,0) rectangle (0,1ex);}
\end{overpic}
\vspace{-1em}
\caption{\label{fig:deltanumdist}
Circular workspace: the worst-case path length occurs when particles must swap antipodes. This can never be achieved but can be asymptotically approached. Plot shows decreasing error (radius $\delta$ around goal positions) as the number of moves grows.
 Red fit line is $8.66/(\textrm{distance}^3)$, which has an R-squared value of 0.77.
} 
\end{figure}

\begin{figure}
\centering
\renewcommand{\figwid}{1\columnwidth}
{
\begin{overpic}[width =\figwid]{contourDistnew.png}\put(-2,10){\begin{turn}{90} \tiny{unique particles}
\end{turn}}
%
\end{overpic}
\vspace{1em}
\begin{overpic}[width =\figwid]{JustSimulationV6.png}\put(-2,10){\begin{turn}{90} \tiny{unique particles}
\end{turn}}
%
\end{overpic}
\begin{overpic}[width =\figwid]{identical.png}\put(-2,6){\begin{turn}{90} \tiny{interchangeable particles}
\end{turn}}
\end{overpic}
}\caption{\label{fig:contourPlots}{Square workspace: starting positions of particles $1$ and $2$ 
 (\protect\tikz \protect\draw[blue,fill=white,line width=0.3mm] (-1ex,0) rectangle (0,1ex);,
 \protect\tikz \protect\draw[myMagenta,fill=white,line width=0.3mm] (-1ex,0) rectangle (0,1ex);)
and goal position of particle $2$ (\protect\tikz \protect\draw[myMagenta,fill=white,line width=0.3mm] (0,0) circle (.5ex);) are fixed, and $\epsilon=0.001$.
 The top row of contour plots show the distance if particle $1$'s goal position is varied in $x$ and $y$. The middle row shows the number of moves required for the same configurations. The bottom row shows the same configuration but when the particles are interchangeable.}
\vspace{-1em}
}
\end{figure}
Algorithm \ref{alg:optimalAlg}  was implemented in Mathematica using particles with zero radius. Figure \ref{fig:reachableSet} shows frames of the algorithm in two representative workspaces, square and disk, with two arbitrary starting and goal configurations.
%An online interactive demonstration and source code of the algorithm are available at \cite{Shahrokhi2015mathematicaParticle}.
%  Fig.~\ref{fig:shapeControlMathematica1}  shows  an implementation of this algorithm with robot initial positions represented by hollow squares and final positions by circles. 
 %Dashed lines show the shortest route if robots could be controlled independently, while solid lines show the optimal shortest  path using uniform inputs.

 The contour plots in Fig.~\ref{fig:contour} left show the length of the path for two different settings. The top row considers \{$s_1,s_2,g_1$\} = \{$(0.2,0.2),(-0.1,-0.1),(0,0)$\} and the bottom row considers  \{$s_1,s_2,g_1$\} = \{$(0.2,0.2),(-0.1,-0.1),(-0.2,0)$\}, each in a workspace with $r= 0.5$, and $g_2$ ranging over all the workspace. Fig.~\ref{fig:contour} left shows the number of moves and right shows the total distance of the path. This plot shows the nonlinear nature of the path planning. When the goal is in the middle of the workspace, a symmetry in the path length is expected as the top row shows. The bottom row shows a shift in the goal position which breaks the symmetry of the path length in the workspace.

### 2018-12-11 journalWallFriction/theory.tex
cleaned images, proofed test, and prepped letter

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Boundary Interaction Model}
\label{sec:theory}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 In the absence of obstacles, uniform inputs move a swarm identically.  
 Independent control requires breaking this symmetry. 
The following sections examine using non-slip boundary contacts to break the symmetry caused by uniform inputs.  
 Our algorithms rely on holding one particle stationary by pushing it into the boundary while moving the other particle. 
 These system dynamics can represent particle swarms in low-Reynolds number environments, where viscosity dominates inertial forces and so velocity is proportional to input force~\cite{Purcell1977}. 
 In this regime, the input force command $\mathbf{u}(t)$ controls the velocity of the particles.  
 If the $i^{\textrm{th}}$ particle has position $\mathbf{x}_i(t)$ and velocity $\dot{\mathbf{x}}_i(t)$,  we assume the following system model:
  \begin{align}\label{eq:swarmDynamicsAndFric} 
\dot{\mathbf{x}}_i(t)
 &= \begin{cases}
  0 &\begin{matrix} \mathbf{x}_i(t) \in  \textrm{boundary \textbf{and}}\\
\mathbf{N}(\textrm{boundary$_{\mathbf{x}_i(t)}$})\cdot   \mathbf{u}(t) \le 0 \end{matrix}
\textrm{    ,} \\
 \mathbf{u}(t) & \textrm{else}.
 \end{cases}
 \end{align}
 Here
 $\mathbf{N}(\textrm{boundary$_{\mathbf{x}_i(t)}$})$ is the normal to the boundary at position $\mathbf{x}_i(t)$ and the frictional force provided by the boundary cancels any control force $ \mathbf{u}(t)$ that pushes into the boundary.

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Boundary Interaction Model}
\label{sec:theory}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 In the absence of obstacles, uniform inputs move a swarm identically.  
 Independent control requires breaking this symmetry. 
The following sections examine using non-slip boundary contacts to break the symmetry caused by uniform inputs.  
 Our algorithms rely on holding one particle stationary by pushing it into the boundary while moving the other particle. 
 These system dynamics can represent particle swarms in low-Reynolds number environments, where viscosity dominates inertial forces and so velocity is proportional to input force~\cite{Purcell1977}. 
 In this regime, the input force command $\mathbf{u}(t)$ controls the velocity of the particles.  
 If the $i^{\textrm{th}}$ particle has position $p_i(t)$ and velocity $\dot{p}_i(t)$,  we assume the following system model:
  \begin{align}\label{eq:swarmDynamicsAndFric} 
\dot{p}_i(t)
 &= \begin{cases}
  0 &\begin{matrix} p_i(t) \in  \textrm{boundary \textbf{and}}\\
\mathbf{N}(\textrm{boundary$_{p_i(t)}$})\cdot   \mathbf{u}(t) \le 0 \end{matrix} \\
 \mathbf{u}(t) & \textrm{else}
 \end{cases}.
 \end{align}
 Here
 $\mathbf{N}(\textrm{boundary$_{p_i(t)}$})$ is the normal to the boundary at position $p_i(t)$ and the frictional force provided by the boundary cancels any control force $ \mathbf{u}(t)$ that pushes into the boundary.
